---
title: "Use of Elsevier Journals by Domain"
author: "UVA StatLab/RDS"
output: html_document
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)
```

```{r setup}
# SETUP ----
# ----------------------------------------------
# Libraries, data, data prep

library(tidyverse)
library(scales)
library(readxl)
library(RColorBrewer)

# read in original unedited data
journal <- read_excel("1science/1figr_U_Virginia_Original.xlsx", 
                      sheet = 4, 
                      skip = 8)

# rename variables in journal
names(journal) <- c("sort", "journal", "issn", "provider", "type", "fig1", "jr1", "jr5", 
                    "cites", "pubs", "synth", "jtier", "subtier", "dupes",
                    "oa", "oaper", "fig2", "scopus", "arif", "domain", "field", "subfield", 
                    "pub2008", "pub2009", "pub2010", "pub2011", "pub2012", "pub2013", "pub2014",
                    "pub2015", "pub2016", "pub2017", "cite2008", "cite2009", "cite2010", "cite2011",
                    "cite2012", "cite2013", "cite2014", "cite2015", "cite2016", "cite2017", 
                    "oa2008", "oa2009", "oa2010", "oa2011", "oa2012", "oa2013", "oa2014", "oa2015", 
                    "oa2016", "oa2017", "oaper2008", "oaper2009", "oaper2010", "oaper2011", "oaper2012",
                    "oaper2013", "oaper2014", "oaper2015", "oaper2016", "oaper2017", "scopus2008",
                    "scopus2009", "scopus2010", "scopus2011", "scopus2012", "scopus2013", "scopus2014",
                    "scopus2015", "scopus2016", "scopus2017", "jr12015", "jr12016", "jr12017", "jr12018",
                    "jid", "journalname")


# derive Current Year Use Percentage - currentper
# derive % of papers referenced by UVA - citeper
# derive UVA Papers % of total - pubsper

journal <- journal %>% 
  mutate(jr5 = if_else(jr5 == "N/A", NA_character_, jr5),
         jr1 = if_else(jr1 == "N/A", NA_character_, jr1),
         currentper = as.numeric(jr5)/as.numeric(jr1),
         citeper = if_else(!is.finite(cites/scopus),0,cites/scopus),
         pubsper = if_else(!is.finite(pubs/scopus),0,pubs/scopus)) 

# derive Downloads JR1 2015-2018
journal <- journal %>% 
  mutate_if(str_detect(names(journal), pattern = "^jr120"), 
            ~as.numeric(if_else(. == "N/A", "0", .))) %>% 
  mutate(jr12015_2018 = jr12015 + jr12016 + jr12017 + jr12018) %>% 
  select(sort:jr5, currentper, cites, citeper, pubs, pubsper,
         synth:jr12018,jr12015_2018,jid, journalname)
```

## Use by Domain

Use of a provider's journal differs by research domain. The following figures show the use -- all article downloads in 2017, current year article downloads in 2017, UVA references, and UVA-authored publications -- by research domain as provided in the 1figr data. Note that this captures the assigned research domain of the article, not of the researcher downloading, referencing, or publishing.


```{r derive}
# USE OF ELSEVIER BY DOMAIN ----
# ----------------------------------------------
# keep only elsevier
els <- journal %>% 
  select(journal, provider, type, jr1, jr5, cites, pubs, synth, scopus, domain, field, jid) %>% 
  filter(type == "Journal" & provider == "Elsevier") %>% 
  mutate(jr1 = as.integer(jr1), jr5 = as.integer(jr5))

# group/summarize by domain
els_group <- els %>% 
  group_by(domain) %>% 
  summarize(titles = n(),
            titlesjr1 = sum(!is.na(jr1)),
            titlesjr5 = sum(!is.na(jr5)),
            jr1 = sum(jr1, na.rm = TRUE),
            jr5 = sum(jr5, na.rm = TRUE),
            refs = sum(cites, na.rm = TRUE),
            pubs = sum(pubs, na.rm = TRUE)) %>% 
  mutate(domain = factor(domain))

# Percents
els_total <- els_group %>% 
  mutate(refs = as.integer(refs),
         pubs = as.integer(pubs)) %>% 
  summarize_if(is.integer, sum)

percentsjr1 <- els_group$jr1/els_total$jr1
percentsjr5 <- els_group$jr5/els_total$jr5
percentsref <- els_group$refs/els_total$refs
percentspub <- els_group$pubs/els_total$pubs
```


### 7a. All Article Downloads by Domain

```{r}
# plot jr1 by domain
els_group %>% 
  mutate(domain = fct_reorder(domain, jr1)) %>% 
  ggplot(aes(x = domain, y = jr1, color = domain)) + 
  geom_segment(aes(x = domain, xend = domain, y = 0, yend = jr1), color = "grey", size = 1) +
  geom_point(size = 5) +
  scale_color_brewer(palette = "BuPu", direction = 1) +
  labs(title = "All 2017 Downloads from Elsevier Journals",
       subtitle = "By Article Domain",
       x = "", y = "Number of Downloads") +
  geom_text(aes(label = comma(jr1)), nudge_x = -.25, color = "black", size = 3) +
  theme(legend.position = "none") +
  coord_flip() 
```

The vast majority of use of Elsevier titles at UVA, whether measured by downloads, citations, or publications, is of articles from the health sciences. For all articles downloaded from Elsevier in 2017, `r percent(percentsjr1[5])` are of articles in the health science domain.


### 7b. Current Year Article Downloads by Domain

```{r}
# plot jr5 by domain
els_group %>% 
  mutate(domain = fct_reorder(domain, jr5)) %>% 
  ggplot(aes(x = domain, y = jr5, color = domain)) + 
  geom_segment(aes(x = domain, xend = domain, y = 0, yend = jr5), color = "grey", size = 1) +
  geom_point(size = 5) +
  scale_color_brewer(palette = "BuPu", direction = 1) +
  labs(title = "Current Year 2017 Downloads from Elsevier Journals",
       subtitle = "By Article Domain",
       x = "", y = "Number of Downloads") +
  geom_text(aes(label = comma(jr5)), nudge_x = -.25, color = "black", size = 3) +
  theme(legend.position = "none") +
  coord_flip() 
```

For current year articles downloaded by UVA researchers from Elsevier in 2017, `r percent(percentsjr5[5])` are of articles in the health science domain.


### 7c. UVA References by Domain

```{r}
# plot refs by domain
els_group %>% 
  mutate(domain = fct_reorder(domain, refs)) %>% 
  ggplot(aes(x = domain, y = refs, color = domain)) + 
  geom_segment(aes(x = domain, xend = domain, y = 0, yend = refs), color = "grey", size = 1) +
  geom_point(size = 5) +
  scale_color_brewer(palette = "BuPu", direction = 1) +
  labs(title = "UVA References by in Elsevier Journals",
       subtitle = "By Article Domain",
       x = "", y = "Number of Downloads") +
  geom_text(aes(label = comma(refs)), nudge_x = -.25, color = "black", size = 3) +
  theme(legend.position = "none") +
  coord_flip() 
```

For references made to articles in Elsevier journals by UVA authors from 2008 to 2017, `r percent(percentsref[5])` are made to articles in the health science domain.


### 7d. UVA-Authored Publications by Domain

```{r}
# plot pubs by domain
els_group %>% 
  mutate(domain = fct_reorder(domain, pubs)) %>% 
  ggplot(aes(x = domain, y = pubs, color = domain)) + 
  geom_segment(aes(x = domain, xend = domain, y = 0, yend = pubs), color = "grey", size = 1) +
  geom_point(size = 5) +
  scale_color_brewer(palette = "BuPu", direction = 1) +
  labs(title = "UVA-Authored Publications in Elsevier Journals",
       subtitle = "By Article Domain",
       x = "", y = "Number of Downloads") +
  geom_text(aes(label = comma(pubs)), nudge_x = -.25, color = "black", size = 3) +
  theme(legend.position = "none") +
  coord_flip() 
```

For articles published in Elsevier journals from 2008 to 2017 by UVA authors, `r percent(percentspub[5])` are made to articles in the health science domain.
