---
title: "Cost per Download for Big 5 Providers"
author: "UVA StatLab/RDS"
output: html_document
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)
```

```{r setup}
library(tidyverse)
library(scales)
library(readxl)

journal <- readRDS("../1science/journal.rds")

# read in supplemental cost data
cost <- read_excel("../1science/1figr_U_Virginia_edit_Supp_Data.xlsx")
names(cost) <- c("provider", "type", "source", "cost", "subcost")
```


## Cost per Download 2017

Adding in information about the cost to UVA for each of the Big 5 providers in 2017, we can calculate the cost per download, or the package cost in dollars in 2017 divided by the number of downloads in 2017. We do this for both all 2017 downloads (JR1) and current year downloads in 2017 (JR5)

```{r derive}
# COST PER DOWNLOAD ----
# ----------------------------------------------
journal_big5 <- journal %>% 
  filter(type == "Package" & provider %in% c("Springer", "Elsevier", "Wiley", "Sage", "Taylor & Francis")) %>% 
  mutate(els = if_else(provider == "Elsevier", "Yes", "No"),
         jr1 = as.integer(jr1),
         jr5 = as.integer(jr5)) %>% 
  select(journal, provider, jr1, jr5, els) %>% 
  left_join(cost, by = "provider") %>% 
  mutate(cpd_jr1 = cost/jr1,
         cpd_jr5 = cost/jr5)
```

### 3a. Cost per Download, all 2017 downloads

```{r}
# CPD-JR1 for Big 5
journal_big5 %>% 
  mutate(provider = factor(provider),
         provider = fct_reorder(provider, desc(cpd_jr1))) %>% 
  ggplot(aes(x = provider, y = cpd_jr1, color = els)) + 
  geom_segment(aes(x = provider, xend = provider, y = 0, yend = cpd_jr1), color = "grey", size = 1) +
  geom_point(size = 5) +
  scale_y_continuous(labels = scales::dollar) +
  scale_color_manual(values = c("orange", "blue")) +
  labs(title = "2017 Cost Per Download, All Downloads/JR1",
       subtitle = "(Package Cost / # JR1 Downloads)",
       x = "", y = "Dollars") +
  geom_text(aes(label = dollar(round(cpd_jr1, 2))), nudge_x = -.3, color = "black", size = 3) +
  theme(legend.position = "none") +
  coord_flip() 
```

This calculation is frequently cited by vendors as “cost-per-use.” By this metric, Elsevier has a cost-per-use of $`r round(journal_big5 %>% filter(provider == "Elsevier") %>% select(cpd_jr1), 2)`. However, this may be misleading insofar as it includes back year content for which we have already purchased perpetual access. 

### 3b. Cost per Download, current year 2017 downloads

```{r}
# CPD-JR5 for Big 5
journal_big5 %>% 
  mutate(provider = factor(provider),
         provider = fct_reorder(provider, desc(cpd_jr5))) %>% 
  ggplot(aes(x = provider, y = cpd_jr5, color = els)) + 
  geom_segment(aes(x = provider, xend = provider, y = 0, yend = cpd_jr5), color = "grey", size = 1) +
  geom_point(size = 5) +
  scale_y_continuous(labels = scales::dollar) +
  scale_color_manual(values = c("orange", "blue")) +
  labs(title = "2017 Cost Per Download, Current Year Downloads/JR5",
       subtitle = "(Package Cost / # JR5 Downloads)",
       x = "", y = "Dollars") +
  geom_text(aes(label = dollar(round(cpd_jr5, 2))), nudge_x = -.3, color = "black", size = 3) +
  theme(legend.position = "none") +
  coord_flip() 
```

Using only current-year downloads generates a higher cost-per-use estimate. This value may be more meaningful, as it reflects only the use of new content added by the current year’s subscription.

## Elsevier Freedom versus Elsevier Subscribed

Same plots as above but with Elsevier broken down by Freedom collection and Subscribed collection. The Freedom Collection is a list of ~1850 active titles (2350 titles in all, including renamed, transferred, discontinued and forthcoming titles) that any institution subscribing to Science Direct Complete gets access to.

### 3c. Cost per Download, all 2017 downloads

```{r}
# Need to derive package level numbers for Elsevier Freedom and Subscribed
tmp <- journal %>% 
  filter(provider == "Elsevier" & type == "Journal") %>% 
  group_by(provider2) %>% 
  summarise(jr1 = sum(as.integer(jr1), na.rm = T),
            jr5 = sum(as.integer(jr5), na.rm = T)) %>% 
  rename(provider = provider2) %>% 
  mutate(els = "Yes") %>% 
  left_join(cost, by = "provider") %>% 
  mutate(cpd_jr1 = cost/jr1,
         cpd_jr5 = cost/jr5)

# bind rows to journal_big5
journal_big6 <- journal_big5 %>% bind_rows(tmp) %>% 
  filter(provider != "Elsevier")


# CPD-JR1 for Big 6
journal_big6 %>% 
  mutate(provider = factor(provider),
         provider = fct_reorder(provider, desc(cpd_jr1))) %>% 
  ggplot(aes(x = provider, y = cpd_jr1, color = els)) + 
  geom_segment(aes(x = provider, xend = provider, y = 0, yend = cpd_jr1), color = "grey", size = 1) +
  geom_point(size = 5) +
  scale_y_continuous(labels = scales::dollar) +
  scale_color_manual(values = c("orange", "blue")) +
  labs(title = "2017 Cost Per Download, All Downloads/JR1",
       subtitle = "(Package Cost / # JR1 Downloads)",
       x = "", y = "Dollars") +
  geom_text(aes(label = dollar(round(cpd_jr1, 2))), nudge_x = -.3, color = "black", size = 3) +
  theme(legend.position = "none") +
  coord_flip() 



```



### 3d. Cost per Download, current year 2017 downloads

```{r}
journal_big6 %>% 
  mutate(provider = factor(provider),
         provider = fct_reorder(provider, desc(cpd_jr5))) %>% 
  ggplot(aes(x = provider, y = cpd_jr5, color = els)) + 
  geom_segment(aes(x = provider, xend = provider, y = 0, yend = cpd_jr5), color = "grey", size = 1) +
  geom_point(size = 5) +
  scale_y_continuous(labels = scales::dollar) +
  scale_color_manual(values = c("orange", "blue")) +
  labs(title = "2017 Cost Per Download, Current Year Downloads/JR5",
       subtitle = "(Package Cost / # JR5 Downloads)",
       x = "", y = "Dollars") +
  geom_text(aes(label = dollar(round(cpd_jr5, 2))), nudge_x = -.3, color = "black", size = 3) +
  theme(legend.position = "none") +
  coord_flip() 

# Figure with just Els Free and Subscribed
journal_big6 %>% 
  mutate(provider = factor(provider),
         provider = fct_reorder(provider, desc(cpd_jr5))) %>% 
  filter(str_detect(provider, "Elsevier")) %>% 
  ggplot(aes(x = provider, y = cpd_jr5)) + 
  geom_segment(aes(x = provider, xend = provider, y = 0, yend = cpd_jr5), color = "grey", size = 1) +
  geom_point(size = 5, color = "blue") +
  scale_y_continuous(labels = scales::dollar) +
  scale_color_manual(values = c("orange", "blue")) +
  labs(title = "2017 Cost Per Download, Current Year Downloads/JR5",
       subtitle = "(Package Cost / # JR5 Downloads)",
       x = "", y = "Dollars") +
  geom_text(aes(label = dollar(round(cpd_jr5, 2))), nudge_x = -.3, color = "black", size = 3) +
  theme(legend.position = "none") +
  coord_flip() 
```

