---
title: "References, Publications, OA-Available over Time"
author: "UVA StatLab/RDS"
output: html_document
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE, fig.width=9)
```

```{r setup}
library(tidyverse)
library(scales)
library(readxl)
library(RColorBrewer)
library(gridExtra)

journal <- readRDS("../1science/journal.rds")

# keep only package data for big 5
big5 <- journal %>% 
  select(provider, type, pub2008:jr12018) %>% 
  filter(type == "Package" & provider %in% c("Springer", "Elsevier", "Wiley", "Sage", "Taylor & Francis"))

# keep only package data for big 5
big5 <- journal %>% 
  # select(provider, type, pub2008:jr12018) %>% 
  select(provider, type, pub2008:oa2017, scopus2008:scopus2017) %>% 
  filter(type == "Package" & provider %in% c("Springer", "Elsevier", "Wiley", "Sage", "Taylor & Francis")) %>% 
  select(-type)

# create a big 6 data frame (two Elseviers: Freedom and Subscribed)
els2 <- journal %>% 
  # select(provider, type, pub2008:jr12018) %>% 
  select(provider2, type, pub2008:oa2017, scopus2008:scopus2017) %>% 
  filter(type == "Journal" & provider2 %in% c("Elsevier Freedom", "Elsevier Subscribed",
                                              "Elsevier Unmatched")) %>% 
  group_by(provider2) %>% 
  summarise_if(is.double, sum, na.rm = TRUE) %>% 
  rename(provider = provider2)

big6 <- bind_rows(big5, els2) %>% 
  filter(provider != "Elsevier") %>% 
  mutate(elsevier = if_else(str_detect(provider, "Elsevier"), 1, 0)) %>% 
  select(provider, elsevier, everything())

rm(els2)


# function to grab legend to use in common
get_legend<-function(myggplot){
  tmp <- ggplot_gtable(ggplot_build(myggplot))
  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
  legend <- tmp$grobs[[leg]]
  return(legend)
}

# function to generate 3 plots + legend
plot_grid <- function(p1, p2, p3){
  legend <- get_legend(p2)
  p1 <- p1 + theme(legend.position="none")
  p2 <- p2 + theme(legend.position="none")
  p3 <- p3 + theme(legend.position="none")
  grid.arrange(p1, p2, p3, legend, ncol = 2)
}

# color palette with 2 reds for Elsevier Free and Sub
# #E41A1C - red for Elsevier
pal <- c("#E41A1C", brewer.pal(n = 5, name = "Set1"))
pal <- c(pal[1:2], "#000000", pal[3:6])

```

## UVA References, UVA Publications, and OA-Available Articles, 2008-2017

### 4a. UVA References in Big 5 Provider Journals over Time
```{r}
# 4. Number of uva citations in Big 5 as percent of citations over time
# reshape time data to long format: pub2008-pub2017, tot2008-tot2017
big5ref <- big5 %>% 
  select(provider, cite2008:cite2017) %>% 
  pivot_longer(-provider, 
               names_to = c(".value", "year"),
               names_sep = -4)

# create total citations
totalref <- journal %>% 
  filter(type == "Package") %>%    # all packages!
  mutate(tot2008 = sum(cite2008), tot2009 = sum(cite2009),
         tot2010 = sum(cite2010), tot2011 = sum(cite2011),
         tot2012 = sum(cite2012), tot2013 = sum(cite2013),
         tot2014 = sum(cite2014), tot2015 = sum(cite2015),
         tot2016 = sum(cite2016), tot2017 = sum(cite2017)) %>% 
  filter(provider == "Elsevier") %>%  # just get the first replication of the totals
  select(tot2008:tot2017) %>% 
  pivot_longer(tot2008:tot2017, 
               names_to = c(".value", "year"),
               names_sep = -4)

# join totalref and create uva percent and indicator for Elsevier
big5ref <- full_join(big5ref, totalref, by = "year") %>% 
  mutate(uvaper = (cite/tot)*100)


# plot uva percent
p1 <- ggplot(totalref, aes(x = year, y = tot, group = 1)) + 
  geom_line() +
  scale_y_continuous(labels = scales::comma) + 
  labs(title = "References by UVA Authors",
       subtitle = "Across All Providers",
       x = "Year", y = "Number of References")

p2 <- ggplot(big5ref, aes(x = year, y = cite, group = provider)) +
  geom_line(aes(color = provider)) +
  scale_color_brewer("Provider", type = "qual", palette = "Set1") +
  labs(title = "References by UVA Authors",
       subtitle = "In Big 5 Providers",
       x = "Year", y = "Number of References")

p3 <- ggplot(big5ref, aes(x = year, y = uvaper, group = provider)) +
  geom_line(aes(color = provider)) +
  scale_color_brewer("Provider", type = "qual", palette = "Set1") +
  labs(title = "Percent of All References by UVA Authors",
       subtitle = "In Big 5 Providers",
       x = "Year", y = "Percent of References")

# create plot
plot_grid(p1, p2, p3)
```

Figure 4 provides trends in citations to the Big 5 providers by UVA authors over time. The top panels reveal an overall growth in research output by UVA researchers, and therefore in citations. The lower panel, showing citations to each provider as a percentage of total citations by UVA authors, suggests that citations to Elsevier titles represent a relatively larger portion of citations relative to the others. This proportion has been flat (if not very slightly in decline) for most of the decade.

### 4b. UVA References in Big 5 Provider Journals over Time

Same figure as above with Elsevier broken down by Freedom versus Subscribed.

```{r }
# Big 6 (Elsevier Free and Elsevier Subscribed)
  big6ref <- big6 %>% 
    select(provider, cite2008:cite2017) %>% 
    pivot_longer(-provider, 
                 names_to = c(".value", "year"),
                 names_sep = -4) %>% 
  full_join(totalref, by = "year") %>% 
  mutate(uvaper = (cite/tot)*100,
         els = if_else(str_detect(provider,"Freedom"), "Yes", "No"))


# plot uva percent
p2 <- ggplot(big6ref, aes(x = year, y = cite, group = provider)) +
  geom_line(aes(color = provider, linetype = provider)) +
  scale_color_manual("Provider", values = pal) +
  scale_linetype_manual("Provider", values = c(2,1,1,1,1,1,1)) +
  labs(title = "References by UVA Authors",
       subtitle = "In Big 5 Providers",
       x = "Year", y = "Number of References")

p3 <- ggplot(big6ref, aes(x = year, y = uvaper, group = provider)) +
  geom_line(aes(color = provider, linetype = provider)) +
  scale_color_manual("Provider", values = pal) +
  scale_linetype_manual("Provider", values = c(2,1,1,1,1,1,1)) +
  labs(title = "Percent of All References by UVA Authors",
       subtitle = "In Big 5 Providers",
       x = "Year", y = "Percent of References")

# create plot
plot_grid(p1, p2, p3)
```



### 5a. UVA Publications in Big 5 Provider Journals over Time
```{r }
# 5. Number of uva-authored publications and percent of uva-authored publications in Big 5 over time
# reshape time data to long format: pub2008-pub2017, scopus2008-scopus2017
big5pub <- big5 %>% 
  select(provider, pub2008:pub2017, scopus2008:scopus2017) %>% 
  pivot_longer(-provider, 
               names_to = c(".value", "year"),
               names_sep = -4) %>% 
  mutate(uvaper = (pub/scopus)*100)

# Big 6 (Elsevier Free and Elsevier Subscribed)
big6pub <- big6 %>% 
  select(provider, pub2008:pub2017, scopus2008:scopus2017) %>% 
  pivot_longer(-provider, 
               names_to = c(".value", "year"),
               names_sep = -4) %>%  
  mutate(uvaper = (pub/scopus)*100,
         els = if_else(str_detect(provider,"Freedom"), "Yes", "No"))



# define plot function
plot_fn <- function(data, y, title, ylab, big6 = FALSE){
  y <- enquo(y)
  p <- ggplot(data, aes(x = year, y = !! y, group = provider))
    if(!big6) {
      p <- p + geom_line(aes(color = provider)) +
        scale_color_brewer(type = "qual", palette = "Set1")
    } else {
      p <- p + geom_line(aes(color = provider, linetype = provider)) +
        scale_color_manual("Provider", values = pal) +
        scale_linetype_manual("Provider", values = c(2,1,1,1,1,1,1))
    } 
  p + labs(title = title,
           subtitle = "In Big 5 Providers",
           x = "Year", y = ylab, color = "Provider")
}
  
p1 <- plot_fn(big5pub, y = scopus, ylab = "Number of Articles",
             title = "Total Number of Articles Published") + 
  scale_y_continuous(labels = scales::comma) 
p2 <- plot_fn(big5pub, y = pub, ylab = "Number of Articles",
             title = "Number of Articles by UVA Authors")
p3 <- plot_fn(big5pub, y = uvaper, ylab = "Percent of Articles",
             title = "Percent of All Articles by UVA Authors")

# create big 5 plot
plot_grid(p1, p2, p3)
```

The top panel indicates that Elsevier and Springer have been publishing more articles over time; at the same time, UVA authors are publishing more articles in the journals provided by Elsevier and Spring. Consequently, the relative amount of UVA work published in Elsevier journals isstable over the last decade. 


### 5b. UVA Publications in Big 5 Provider Journals over Time

Same figure as above with Elsevier broken down by Freedom versus Subscribed.

```{r }
p1 <- plot_fn(big6pub, y = scopus, ylab = "Number of Articles",
             title = "Total Number of Articles Published", 
             big6 = TRUE) + 
  scale_y_continuous(labels = scales::comma) 
p2 <- plot_fn(big6pub, y = pub, ylab = "Number of Articles",
             title = "Number of Articles by UVA Authors", 
             big6 = TRUE)
p3 <- plot_fn(big6pub, y = uvaper, ylab = "Percent of Articles",
             title = "Percent of All Articles by UVA Authors", 
             big6 = TRUE)

# create big 6 plot
plot_grid(p1, p2, p3)

```


### 6a. OA-Available Articles in Big 5 Provider Journals over Time

```{r }
# 6. Number of oa-available publications and percent of oa-available publications in Big 5 over time
# reshape time data to long format: oa2008-ao2017, scopus2008-scopus2017
big5oa <- big5 %>% 
  select(provider, oa2008:oa2017, scopus2008:scopus2017) %>% 
  pivot_longer(-provider, 
               names_to = c(".value", "year"),
               names_sep = -4) %>% 
  mutate(oaper = (oa/scopus)*100)

big6oa <- big6 %>% 
  select(provider, oa2008:oa2017, scopus2008:scopus2017) %>% 
  pivot_longer(-provider, 
               names_to = c(".value", "year"),
               names_sep = -4) %>% 
  mutate(oaper = (oa/scopus)*100,
         els = if_else(str_detect(provider,"Freedom"), "Yes", "No"))


# labels to add 2017 values at end of lines - big 5
p3_labels <- big5oa %>% 
  group_by(provider) %>% 
  filter(year == "2017") %>% 
  mutate(oaper = round(oaper))

# labels to add 2017 values at end of lines - big 6
p3_labels2 <- big6oa %>% 
  group_by(provider) %>% 
  filter(year == "2017") %>% 
  mutate(oaper = round(oaper))


p1 <- plot_fn(big5oa, y = scopus, ylab = "Number of Articles",
              title = "Number of Articles Published") + 
  scale_y_continuous(labels = scales::comma) 
p2 <- plot_fn(big5oa, y = oa, ylab = "Number of Articles",
              title = "Number of OA-Available Articles")
p3 <- plot_fn(big5oa, y = oaper, ylab = "Percent of Articles",
              title = "Percent of OA-Available Articles") +
  geom_text(aes(x = year, y = oaper, label = oaper), p3_labels, nudge_x = 0.2)

# create big 5 plot
plot_grid(p1, p2, p3)
```

The top panel, again, indicates that Elsevier and Springer have been publishing more articles over time, and the number of OA-available articles in these providers' titles has been increasing as well. The drop off in OA availability for more recent content likely reflects the rolling embargo window imposed by vendors. Piwowar et al. found in a study of national OA availability trends that OA availability is increasing in recent years, and that for the most recent year they studied, 2015, 45% of the articles were available in OA versions. However, it seems that this percentage varies strongly depending on the provider, with Elsevier and Taylor & Francis keeping more of their content behind paywalls.

### 6b. OA-Available Articles in Big 5 Provider Journals over Time

Same figure as above with Elsevier broken down by Freedom versus Subscribed.

```{r }
p1 <- plot_fn(big6oa, y = scopus, ylab = "Number of Articles",
              title = "Number of Articles Published", 
              big6 = TRUE) + 
  scale_y_continuous(labels = scales::comma) 
p2 <- plot_fn(big6oa, y = oa, ylab = "Number of Articles",
              title = "Number of OA-Available Articles", 
              big6 = TRUE)
p3 <- plot_fn(big6oa, y = oaper, ylab = "Percent of Articles",
              title = "Percent of OA-Available Articles", 
              big6 = TRUE) +
  geom_text(aes(x = year, y = oaper, label = oaper), p3_labels2, nudge_x = 0.2, size = 3)

# create big 6 plot
plot_grid(p1, p2, p3)
```

