---
title: "Cost per Download for Big 5 Providers"
author: "UVA StatLab/RDS"
output: html_document
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)
```

```{r setup}
library(tidyverse)
library(scales)
library(readxl)

# read in original unedited data
journal <- read_excel("../1science/1figr_U_Virginia_Original.xlsx", 
                      sheet = 4, 
                      skip = 8)

# rename variables in journal
names(journal) <- c("sort", "journal", "issn", "provider", "type", "fig1", "jr1", "jr5", 
                    "cites", "pubs", "synth", "jtier", "subtier", "dupes",
                    "oa", "oaper", "fig2", "scopus", "arif", "domain", "field", "subfield", 
                    "pub2008", "pub2009", "pub2010", "pub2011", "pub2012", "pub2013", "pub2014",
                    "pub2015", "pub2016", "pub2017", "cite2008", "cite2009", "cite2010", "cite2011",
                    "cite2012", "cite2013", "cite2014", "cite2015", "cite2016", "cite2017", 
                    "oa2008", "oa2009", "oa2010", "oa2011", "oa2012", "oa2013", "oa2014", "oa2015", 
                    "oa2016", "oa2017", "oaper2008", "oaper2009", "oaper2010", "oaper2011", "oaper2012",
                    "oaper2013", "oaper2014", "oaper2015", "oaper2016", "oaper2017", "scopus2008",
                    "scopus2009", "scopus2010", "scopus2011", "scopus2012", "scopus2013", "scopus2014",
                    "scopus2015", "scopus2016", "scopus2017", "jr12015", "jr12016", "jr12017", "jr12018",
                    "jid", "journalname")


# derive Current Year Use Percentage - currentper
# derive % of papers referenced by UVA - citeper
# derive UVA Papers % of total - pubsper

journal <- journal %>% 
  mutate(jr5 = if_else(jr5 == "N/A", NA_character_, jr5),
         jr1 = if_else(jr1 == "N/A", NA_character_, jr1),
         currentper = as.numeric(jr5)/as.numeric(jr1),
         citeper = if_else(!is.finite(cites/scopus),0,cites/scopus),
         pubsper = if_else(!is.finite(pubs/scopus),0,pubs/scopus)) 

# derive Downloads JR1 2015-2018
journal <- journal %>% 
  mutate_if(str_detect(names(journal), pattern = "^jr120"), 
            ~as.numeric(if_else(. == "N/A", "0", .))) %>% 
  mutate(jr12015_2018 = jr12015 + jr12016 + jr12017 + jr12018) %>% 
  select(sort:jr5, currentper, cites, citeper, pubs, pubsper,
         synth:jr12018,jr12015_2018,jid, journalname)

# read in supplemental cost data
cost <- read_excel("../1science/1figr_U_Virginia_edit_Supp_Data.xlsx")
names(cost) <- c("provider", "type", "source", "cost", "subcost")
```


## Cost per Download 2017

Adding in information about the cost to UVA for each of the Big 5 providers in 2017, we can calculate the cost per download, or the package cost in dollars in 2017 divided by the number of downloads in 2017. We do this for both all 2017 downloads (JR1) and current year downloads in 2017 (JR5)

```{r derive}
# COST PER DOWNLOAD ----
# ----------------------------------------------
journal_big5 <- journal %>% 
  filter(type == "Package" & provider %in% c("Springer", "Elsevier", "Wiley", "Sage", "Taylor & Francis")) %>% 
  mutate(els = if_else(provider == "Elsevier", "Yes", "No"),
         jr1 = as.integer(jr1),
         jr5 = as.integer(jr5)) %>% 
  select(journal, provider, jr1, jr5, els) %>% 
  left_join(cost, by = "provider") %>% 
  mutate(cpd_jr1 = cost/jr1,
         cpd_jr5 = cost/jr5)
```

### 3a. Cost per Download, all 2017 downloads

```{r}
# CPD-JR1 for Big 5
journal_big5 %>% 
  mutate(provider = factor(provider),
         provider = fct_reorder(provider, desc(cpd_jr1))) %>% 
  ggplot(aes(x = provider, y = cpd_jr1, color = els)) + 
  geom_segment(aes(x = provider, xend = provider, y = 0, yend = cpd_jr1), color = "grey", size = 1) +
  geom_point(size = 5) +
  scale_y_continuous(labels = scales::dollar) +
  scale_color_manual(values = c("orange", "blue")) +
  labs(title = "2017 Cost Per Download, All Downloads/JR1",
       subtitle = "(Package Cost / # JR1 Downloads)",
       x = "", y = "Dollars") +
  geom_text(aes(label = dollar(round(cpd_jr1, 2))), nudge_x = -.3, color = "black", size = 3) +
  theme(legend.position = "none") +
  coord_flip() 
```

This calculation is frequently cited by vendors as “cost-per-use.” By this metric, Elsevier has a cost-per-use of $`r round(journal_big5 %>% filter(provider == "Elsevier") %>% select(cpd_jr1), 2)`. However, this may be misleading insofar as it includes back year content for which we have already purchased perpetual access. 

### 3b. Cost per Download, current year 2017 downloads

```{r}
# CPD-JR5 for Big 5
journal_big5 %>% 
  mutate(provider = factor(provider),
         provider = fct_reorder(provider, desc(cpd_jr5))) %>% 
  ggplot(aes(x = provider, y = cpd_jr5, color = els)) + 
  geom_segment(aes(x = provider, xend = provider, y = 0, yend = cpd_jr5), color = "grey", size = 1) +
  geom_point(size = 5) +
  scale_y_continuous(labels = scales::dollar) +
  scale_color_manual(values = c("orange", "blue")) +
  labs(title = "2017 Cost Per Download, Current Year Downloads/JR5",
       subtitle = "(Package Cost / # JR5 Downloads)",
       x = "", y = "Dollars") +
  geom_text(aes(label = dollar(round(cpd_jr5, 2))), nudge_x = -.3, color = "black", size = 3) +
  theme(legend.position = "none") +
  coord_flip() 
```

Using only current-year downloads generates a higher cost-per-use estimate. This value may be more meaningful, as it reflects only the use of new content added by the current year’s subscription.