---
title: "High Use Journals in Big 5 Providers"
author: "UVA StatLab/RDS"
output: html_document
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)
```

```{r setup}
library(tidyverse)
library(RColorBrewer)
library(scales)

journal <- readRDS("../1science/journal.rds")
  
```


```{r functions}
# function to get high use counts
highuse_fn <- function(group_var){
  group_var <- enquo(group_var)
  
  journal %>% 
    select(journal, provider, provider2, type, jr1, jr5, cites, pubs, synth, scopus, domain, field, jid) %>% 
    filter(type == "Journal") %>% 
    mutate(provider = factor(provider),
           big5 = if_else(provider %in% c("Springer", "Elsevier", "Wiley", "Sage", "Taylor & Francis"), "Yes", "No"),
           els = if_else(provider == "Elsevier", "Yes", "No")) %>% 
    mutate(jr1 = as.integer(jr1), jr5 = as.integer(jr5)) %>% 
    group_by(!! group_var) %>% 
    arrange(!! group_var, desc(jr1)) %>% 
    mutate(cum_jr1 = cumsum(jr1),
           prov_jr1 = max(cum_jr1, na.rm=T),
           cum_jr1_per = (cum_jr1/prov_jr1)*100,
           use_jr1 = case_when(cum_jr1_per <= 80 ~ "JR80 Titles",
                               cum_jr1_per > 80 & cum_jr1_per <= 90 ~ "JR90 Titles",
                               cum_jr1_per > 90 & cum_jr1_per <= 95 ~ "JR95 Titles",
                               TRUE ~ "Remaining Titles")) %>% 
    arrange(!! group_var, desc(jr5)) %>%
    mutate(cum_jr5 = cumsum(jr5),
           prov_jr5 = max(cum_jr5, na.rm=T),
           cum_jr5_per = (cum_jr5/prov_jr5)*100,
           use_jr5 = case_when(cum_jr5_per <= 80 ~ "JR80 Titles",
                               cum_jr5_per > 80 & cum_jr5_per <= 90 ~ "JR90 Titles",
                               cum_jr5_per > 90 & cum_jr5_per <= 95 ~ "JR95 Titles",
                               TRUE ~ "Remaining Titles")) %>% 
    arrange(!! group_var, desc(cites)) %>% 
    mutate(cum_refs = cumsum(cites),
           prov_refs = max(cum_refs, na.rm=T),
           cum_refs_per = (cum_refs/prov_refs)*100,
           use_refs = case_when(cum_refs_per <= 80 ~ "JR80 Titles",
                                cum_refs_per > 80 & cum_refs_per <= 90 ~ "JR90 Titles",
                                cum_refs_per > 90 & cum_refs_per <= 95 ~ "JR95 Titles",
                                TRUE ~ "Remaining Titles")) %>% 
    arrange(!! group_var, desc(pubs)) %>% 
    mutate(cum_pubs = cumsum(pubs),
           prov_pubs = max(cum_pubs, na.rm=T),
           cum_pubs_per = (cum_pubs/prov_pubs)*100,
           use_pubs = case_when(cum_pubs_per <= 80 ~ "JR80 Titles",
                                cum_pubs_per > 80 & cum_pubs_per <= 90 ~ "JR90 Titles",
                                cum_pubs_per > 90 & cum_pubs_per <= 95 ~ "JR95 Titles",
                                TRUE ~ "Remaining Titles"))  
}

# Function to create figures

fig1 <- function(use, provider, title, subtitle){
  use <- enquo(use)
  provider <- enquo(provider)
  highuse_counts <- highuse %>% ungroup() %>% 
    mutate(!! use := factor(!! use, levels = c("JR80 Titles", "JR90 Titles", "JR95 Titles", "Remaining Titles"))) %>% 
    filter(big5 == "Yes") %>% 
    count(!! provider, !! use) %>% group_by(!! provider) %>% 
    mutate(pct=n/sum(n), ypos = cumsum(pct), num_journal = sum(n))
  
  assign(x = "highuse_counts", highuse_counts, envir = .GlobalEnv)
  
  p <- highuse_counts %>% 
    ggplot(aes(x = !! provider, y = pct, fill = fct_rev(!! use), label = n)) +
    geom_col() +
    scale_y_continuous(labels = scales::percent) +
    scale_fill_brewer(type = 'seq', palette = 'BuPu') +
    labs(title = title,
         subtitle = subtitle,
         x = "", y = "Percent") +
   # geom_text(aes(label=paste0(n), y=ypos - 0.05)) +
    geom_text(aes(label=paste0(num_journal), y = 1.05), size = 3) +
    guides(fill = guide_legend(reverse=TRUE, title = NULL)) +
    theme(legend.position = "bottom") +
    coord_flip()
  plotly::ggplotly(p)
}

```


We began by evaluating the proportion of highly-used titles in 2017 within the Big 5 provider packages: the journals that account for the top X% of use, examined through multiple proxies for “use”, including 

* all article downloads in 2017 (JR1)
* current year article downloads in 2017 (JR5)
* references to journal articles made by UVA authors from 2008-2017
* publications by UVA authors from 2008-2017

For each use metric we graph the percent of titles within each provider's package that are among the top 80% used titles (JR80 Titles), the top 90% used titles (JR90 Titles), and the top 95% used titles (JR95 Titles) at UVA. The number of titles represented in each level of use is printed on the figures as well, along with the number of titles that make up the remaining 5% of use.

```{r derive}
# Generate JR80, JR90, JR95 measures using jr1 (all 2017 downloads), jr5 (current year 2017 downloads), references, publications
highuse <- highuse_fn(group_var = provider)

```

### 1a. High-use Journals by All Article Downloads
```{r}
# plot use_jr1 for big 5
fig1(use_jr1, 
     provider,
     title = "Percentage of Titles in High Use Journals",
     subtitle = "By All Downloads in 2017 (JR1)")
```

Among Elsevier journals, `r highuse_counts[1,3]` or `r sprintf("%1.1f", highuse_counts[1,4]*100)`% of the titles are among the journals that make up 80% of downloads at UVA in 2017. Another `r highuse_counts[2,3]` titles, or `r sprintf("%1.1f", (highuse_counts[1,4] + highuse_counts[2,4])*100)`% overall, are among the journals that make up 90% of downloads at UVA. Including journals that make up 95% of downlaods adds another `r highuse_counts[3,3]` titles from Elsevier, or `r sprintf("%1.1f", (highuse_counts[1,4] + highuse_counts[2,4] + highuse_counts[3,4])*100)`% of the titles in the Elsevier package. Finally, `r highuse_counts[4,3]` journals in Elsevier's package, or `r  sprintf("%1.1f", highuse_counts[4,4]*100)`%, are not among the most highly-used journals at UVA by this metric.

The pattern for Springer is similar to that of Elsevier. Taylor & Francis and Wiley have a slightly higher ratio of high-use journals, while Sage has the highest rate of high-use journals.

### 1b. High-use Journals by Current Year Article Downloads
```{r}
# plot use_jr5 for big 5
fig1(use_jr5, 
     provider,
     title = "Percentage of Titles in High Use Journals",
     subtitle = "By Current Year Downloads in 2017 (JR5)")

```

Looking at only downloads of articles from published in the current year, among Elsevier journals, `r highuse_counts[1,3]` or `r sprintf("%1.1f", highuse_counts[1,4]*100)`% of the titles are among the journals that make up 80% of use at UVA. Another `r highuse_counts[2,3]` titles, or `r sprintf("%1.1f", (highuse_counts[1,4] + highuse_counts[2,4])*100)`% overall, are among the journals that make up 90% of current year downloads within Elsevier; and an additional `r highuse_counts[3,3]` titles from Elsevier, or `r sprintf("%1.1f", (highuse_counts[1,4] + highuse_counts[2,4] + highuse_counts[3,4])*100)`% of the titles in the Elsevier package, account for 95% of use.  `r highuse_counts[4,3]` journals in Elsevier's package, or `r  sprintf("%1.1f", highuse_counts[4,4]*100)`%, are not among the most highly-used journals at UVA by this metric.

Both Springer and Wiley have similar ratios of high-use journals by current year downloads. Taylor & Francis has a lower ratio and Sage shows a higher ratio of highly-used journals by this metric.

### 1c. High-use Journals by References from UVA Authors
```{r}
# plot use_refs for big 5
fig1(use_refs, 
     provider,
     title = "Percentage of Titles in High Use Journals",
     subtitle = "By References by UVA Authors")
```

Measuring highly-used journals by focusing on the journals that UVA researchers have cited in the last 10 years shrinks the pool of titles categorized as high use. Among Elsevier journals, `r highuse_counts[1,3]` or `r sprintf("%1.1f", highuse_counts[1,4]*100)`% of the titles are among the journals that make up 80% of references by researchers at UVA. Using the journals that make up 90% of references adds `r highuse_counts[2,3]` Elsevier titles, or `r sprintf("%1.1f", (highuse_counts[1,4] + highuse_counts[2,4])*100)`% of the package overall. Looking at the journals that make up 95% of references adds an additional `r highuse_counts[3,3]` titles from Elsevier, or `r sprintf("%1.1f", (highuse_counts[1,4] + highuse_counts[2,4] + highuse_counts[3,4])*100)`% of the titles in the Elsevier package. `r highuse_counts[4,3]` journals in Elsevier's package, or `r  sprintf("%1.1f", highuse_counts[4,4]*100)`%, are not among the most highly-used journals at UVA by this metric.

### 1d. High-use Journals by UVA-Authored Publications
```{r}
# plot use_pubs for big 5
fig1(use_pubs, 
     provider,
     title = "Percentage of Titles in High Use Journals",
     subtitle = "By UVA Authored Papers")
```

Focusing on the journals UVA researchers have published in over the last 10 years, `r highuse_counts[1,3]` or `r sprintf("%1.1f", highuse_counts[1,4]*100)`% of Elsevier titles are among the journals that make up 80% of publications by UVA authors. The JR90 measure expands that to an additional `r highuse_counts[2,3]` Elsevier titles, or `r sprintf("%1.1f", (highuse_counts[1,4] + highuse_counts[2,4])*100)`% of the package; the JR95 measure expands that to `r highuse_counts[3,3]` additional Elsevier titles, or `r sprintf("%1.1f", (highuse_counts[1,4] + highuse_counts[2,4] + highuse_counts[3,4])*100)`% of the titles in the Elsevier package. `r highuse_counts[4,3]` journals in Elsevier's package, or `r  sprintf("%1.1f", highuse_counts[4,4]*100)`%, are not among the most highly-used journals at UVA by this metric.

## Elsevier Freedom versus Elsevier Subscribed

The figures above are repeated below but with Elsevier journals disaggregated into the Freedom collection, Subscribed collection, and journals we were not able to match. Here "high use" is calculated within the whole Elsevier package, as above, but visualized by the disaggregated Elsevier sets. 

```{r derive2}
# Big 5 with Elsevier broken down by Subscribed and Freedom
# highuse <- highuse_fn(group_var = provider)
```

### 1e. High-use Journals by All Article Downloads
```{r}
fig1(use_jr1, 
     provider2,
     title = "Percentage of Titles in High Use Journals",
     subtitle = "By All Downloads in 2017 (JR1)")
```


With regard to article downloads, the majority of journals in the Subscribed Collection are high-use journals by the JR80 measure (418 of 726 titles). Only 15% of the titles in the Freedom Collection are highly used by the JR80 measure( 166 of 1086 titles). 

### 1f. High-use Journals by Current Year Article Downloads

```{r}
fig1(use_jr5, 
     provider2,
     title = "Percentage of Titles in High Use Journals",
     subtitle = "By Current Year Downloads in 2017 (JR5)")

```

Focusing on current-year downloads only, 335 journals in the Subscribed Collection are high-use journals by the JR80 metric (46%); 136 journals in the Freedom Collection hit this mark (13%).

### 1g. High-use Journals by References from UVA Authors

```{r}
fig1(use_refs, 
     provider2,
     title = "Percentage of Titles in High Use Journals",
     subtitle = "By References by UVA Authors")

```

Looking at usage based on citations, 306 journals in the Subscribed Collection are high-use by the JR80 metric (42%) compared to 65 journals in the Freedom Collection (6%).

### 1h. High-use Journals by UVA-Authored Publications

```{r}
fig1(use_pubs, 
     provider2,
     title = "Percentage of Titles in High Use Journals",
     subtitle = "By UVA Authored Papers")
```

Finally, with regard to UVA-authored publicatons, 270 journals in the Subscribed Collection are highly-used by the JR80 metric (37%) while 75 journals in the Freedom Collection are high-use (7%).